Notes Dendro Analysis

### 2018-01-16 analysis/rwi/climate_rwi_relate.Rmd
UPDATE

* Correlación entre Marzo del mes anterior y hasta Sep del mes en curso
* Correlación con tmax winter, summer, spring, autumn, hydrol, may-jul, june-july

```{r}
set.seed(3333)
rtmax_SJ <- dcc(chrono = cro_sj, climate = tmax_N, 
         method = "correlation",  boot = "std", 
        .range(-3:9) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11) + .mean(5:7) + .mean(6:7))

rtmax_caH <- dcc(chrono = cro_caH, climate = tmax_S, 
         method = "correlation",  boot = "std", 
         .range(-3:9) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11) + .mean(5:7) + .mean(6:7))

rtmax_caL <- dcc(chrono = cro_caL, climate = tmax_S, 
         method = "correlation",  boot = "std", 
         .range(-3:9) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11) + .mean(5:7) + .mean(6:7))

>>>>

* Correlación entre Marzo del mes anterior y hasta Sep del mes en curso
* Correlación con tmax winter, summer, spring, autumn, hydrol

```{r}
set.seed(3333)
rtmax_SJ <- dcc(chrono = cro_sj, climate = tmax_N, 
         method = "correlation",  boot = "std", 
        .range(-3:9) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

rtmax_caH <- dcc(chrono = cro_caH, climate = tmax_S, 
         method = "correlation",  boot = "std", 
         .range(-3:9) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

rtmax_caL <- dcc(chrono = cro_caL, climate = tmax_S, 
         method = "correlation",  boot = "std", 
         .range(-3:9) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

### 2018-01-16 man/ms_0_intro.Rmd
updating
UPDATE

* To quantify how two extreme drought events influenced the greennes and radial growth of *Q. pyreancia* forests in their rear edge,

* to analyze the resilience of these forests to sucessive extrme drought events,

@@

>>>>

* To quantify how two extreme drought events influenced the greenness and radial growth of *Q. pyreancia* forests in their rear edge,

* to analyze the resilience of these forests to successive extreme drought events,

@@

Do northern and southern populations of Pyrenean oak forests differ in their resistance, resilience or recovery to extreme drought events?

### 2018-01-16 man/ms_1_material_methods.Rmd
UPDATE

---
title: "material and methods"
output:  
  md_document:
    toc: true
    variant: markdown_github
bibliography: ../references.bib
templat
csl: ../geb.csl
---

@@

### Tree sampling 
Samplig was carried during autumn of 2016. Trees were sampled at two locations in contrasting slopes of Sierra Nevada: San Juan (SJ; northern site) and Cáñar (CA; southern site) (Figure 1; Table 1). Two elevations were sampled for the southern site (CA-Low and CA-High). All the sites were oak monospecific and representatives of two of the three the population's cluster identified for the specie in this mountain range (:red_circle: mejorar; citar Pérz-Luque et al..). In each site between 15 and 20 dominant trees were randomly selected. Two cores of 5 mm of diameter were taken from each at breast heith (1.3 m) using an increment borer. Diameter at breast height (DBH) and total height were measured using a girth tape and a Vertex IV ultrasonic hypsometer (Haglöf, Sweden) respectively. We assessed stand competition of the target tree by recording all neighboring living trees with dbh > 7.5 cm in a circle of 10 m radi. The species, dbh and total height were annotated. We also measured distance and azimuth to the target tree. (:red_circle: indices competencia??)

### Dendroecological analyses 
Tree cores were air dried, glued onto wooden mounts and sanded. Annual radial growth (ring width, RW) were measured with a LINTAB measuring device (Rinntech, Heidelberg, Germany) coupled to a stereomicroscope, with an accuracy of 0.01 mm. Individual ring series were visually and statistically cross-dated with TSAP software (Rinntech, Heidelberg, Germany), using the statistics Gleichläufigkeit (GLK), t-value and the crossdating index (CDI). Validation of the croos-dating was done using COFECHA software [@Holmes1983].

The growth trends were analyzed at different time scales. To study the response of growth to the inter-annual variability of climate (short-term response) we used pre-withened residual chronologies calculated from ratios between raw growth measurements and individual cubic splines with a 50 % frecuency cutoff at 30 years [@Fritts1976]. Tree-ring width series were standardized and detrended using the `dplR` [@Bunn2008, @Bunn2010] package. Residual site chronologies were obtained computing the bi-weight robust mean of all prewhitened growth indices for the trees of the same site [@Cook1990]. The statistical quality of each chronology was checked via expressed population signal [@Wigley1984]. A threshold value of EPS > 0.85 was used to determine the cutoff year of the time span that could be considered reliable.

The long-term growth response was analyzed using basal area increment (hereafter BAI, $\mathrm{cm^2 \cdot year^{-1}}$). BAI represents a more accurate indicator of growth than ring-width, since it removes variation in growth attributatle to increasing stem circumference [@Biondi2008]. We used measured dbh and raw nondetrened ring-widths to compute BAI by subtracting twice the annual ring width from the annual diameter, starting from the measured diameter outside the bark [@Piovesan2008]. We used the following equation:

@@

Climate data were obtained from the European Daily High-Resolution Observational Gridded Dataset (E-OBS v16) [@Haylock2008]. Monthly precipitation and mean, minimum and maximum temperatures were obtained at a 0.25 x 0.25 º resolution for the 1950-2016 period. We selected grid cells covering each sampled sites. Data were downloaded and preprocess using the climate4R bundle (http://www.meteo.unican.es/en/climate4R) [@Cofino2018, @Frias2018].
We also used the Standardized Precipitation-Evapotranspiration Index (SPEI), a multiscalar drought index that incorporates both precipitation and temperature [@VicenteSerrano2010, @Begueria2014]. SPEI values for the period 1961-2014 were obtained with a spatial resolution of 1.1 km from the Drought indices dataset for Spain database (http://monitordesequia.csic.es/), a high resolution database of drought indices for Spain [@VicenteSerrano2017]. We select a temporal scale of 6 months.

@@

$$\mathrm{\% GC} = \left [ \frac{\left ( M1 - M2 \right )}{M2}\right] \times 100 $$
where $M1$ is preceding 10-year median and $M2$ is subsequent 10-year median. We used medians since they are mores robust estimator of central tendency than means @Rubino2004 @Camarero2011]. Site disturbance chronologies were constructed by averaging the individual disturbances series annually. A

To separate growth periods produced by disturbance events from those by climate, we considered a treshold of 25 % of GC for consi

>>>>

---
title: "material and methods"
output:  
  md_document:
    toc: true
    variant: markdown_github
bibliography: ../references.bib
csl: ../ecology.csl
---

@@

### Tree sampling 
Sampling was carried during autumn of 2016. Trees were sampled at two locations in contrasting slopes of Sierra Nevada: San Juan (SJ; northern site) and Cáñar (CA; southern site) (Figure 1; Table 1). For the southern site two elevations were sampled: CA-Low and CA-High. All the sites were oak monospecific and representatives of two of the three the population's cluster identified for the specie in this mountain range (:red_circle: mejorar; citar Pérz-Luque et al..). In each site between 15 and 20 dominant trees were randomly selected. Two cores of 5 mm of diameter were taken from each at breast heigth (1.3 m) using an increment borer. Diameter at breast height (DBH) and total height were measured using a girth tape and a Vertex IV ultrasonic hypsometer (Haglöf, Sweden) respectively. We assessed stand competition of the target tree by recording all neighboring living trees with dbh > 7.5 cm in a circle of 10 m radi. The species, dbh and total height were annotated. We also measured distance and azimuth to the target tree. (:red_circle: indices competencia??)

### Dendroecological analyses 
Tree cores were air dried, glued onto wooden mounts and sanded. Annual radial growth (ring width, RW) were measured with a LINTAB measuring device (Rinntech, Heidelberg, Germany) coupled to a stereomicroscope, with an accuracy of 0.01 mm. Individual ring series were visually and statistically cross-dated with TSAP software (Rinntech, Heidelberg, Germany), using the statistics Gleichläufigkeit (GLK), t-value and the crossdating index (CDI). Validation of the cross-dating was done using COFECHA software [@Holmes1983].

The growth trends were analyzed at different time scales. To study the response of growth to the inter-annual variability of climate (short-term response) we used pre-whitened residual chronologies calculated from ratios between raw growth measurements and individual cubic splines with a 50 % frequency cutoff at 30 years [@Fritts1976]. Tree-ring width series were standardized and detrended using the `dplR` [@Bunn2008; @Bunn2010] package. Residual site chronologies were obtained computing the biweight robust mean of all prewhitened growth indices for the trees of the same site [@Cook1990]. The statistical quality of each chronology was checked via expressed population signal [@Wigley1984]. A threshold value of EPS > 0.85 was used to determine the cutoff year of the time span that could be considered reliable.

The long-term growth response was analyzed using basal area increment (hereafter BAI, $\mathrm{cm^2 \cdot year^{-1}}$). BAI represents a more accurate indicator of growth than ring-width, since it removes variation in growth attributable to increasing stem circumference [@Biondi2008]. We used measured dbh and raw nondetrened ring-widths to compute BAI by subtracting twice the annual ring width from the annual diameter, starting from the measured diameter outside the bark [@Piovesan2008]. We used the following equation:

@@

Climate data were obtained from the European Daily High-Resolution Observational Gridded Dataset (E-OBS v16) [@Haylock2008]. Monthly precipitation and mean, minimum and maximum temperatures were obtained at a 0.25 x 0.25 º resolution for the 1950-2016 period. We selected grid cells covering each sampled sites. Data were downloaded and preprocess using the climate4R bundle (http://www.meteo.unican.es/en/climate4R) [@Cofino2018, @Frias2018].
We also used the Standardized Precipitation-Evapotranspiration Index (SPEI), a multiscalar drought index that incorporates both precipitation and temperature [@VicenteSerrano2010; @Begueria2014]. SPEI values for the period 1961-2014 were obtained with a spatial resolution of 1.1 km from the Drought indices dataset for Spain database (http://monitordesequia.csic.es/), a high resolution database of drought indices for Spain [@VicenteSerrano2017]. We select a temporal scale of 6 months.

@@

$$\mathrm{\% GC} = \left [ \frac{\left ( M1 - M2 \right )}{M2}\right] \times 100 $$
where $M1$ is preceding 10-year median and $M2$ is subsequent 10-year median. We used medians since they are mores robust estimator of central tendency than means [@Rubino2004; @Camarero2011]. Site disturbance chronologies were constructed by averaging the individual disturbances series annually. A

To separate growth periods produced by disturbance events from those by climate, we considered a threshold of 25 % of GC for consi

### 2018-01-16 man/ms_3_discussion.Rmd
update ms and add compile document
UPDATE

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

>>>>

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

## Reducción crecimiento y sequías

Se observó una gran reducción de crecimiento en 1994-1995, 1999 y 2005 para Pinus nigra en el SE de Andalusia [@SanchezSalguero2013] coincidiendo con años de sequía para esa zona. La máxima reducción en crecimiento fue observada para 2005, siendo significativamente mayor que para el resto de años [@SanchezSalguero2013].

## Relaciones crecimiento-clima en otras especies

### 2018-01-16 man/notes_dendro_analysis.Rmd
UPDATE

>>>>

# Dendro statistics

see Camarero et al 2013. Ann Forest sciences 
* Mean and standard deviation (SD) of raw tree-ring width data
* First-order autocorrelation of raw data (AC1), which measures the year-to-year persistence in ring width
* Mean sensitivity (MS) of the residual ring-width series, which measures the year-to-year variability in the width index
* Mean between-trees correlation within each site (r bt), which quantifies the similarity in residual width indices among trees
* Expressed population signal (EPS) of residual series, which assesses the representativeness of the sample size compared with theoretical infinite population
